---
title: "Getting and Cleaning Data Course Project"


output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---
Run Analysis : HTML File

By: Raj Kumar Pandit


##Preliminaries


####Packages Initialization

```{r message=FALSE}
packages <- c("data.table", "reshape2")
sapply (packages, require, character.only=TRUE, quietly=TRUE)
library(knitr)
library(markdown)
```

####Set Path.

```{r message=FALSE}
pth <- getwd()
pth
```


##Get the data


Downloading the file. 

```{r, eval=FALSE}
url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
download.file(url, "Dataset.zip" )
```

Unzipping the file- The R way :) 

```{r, eval=FALSE}
executable <- file.path("C:", "Program Files", "7-Zip", "7zFM.exe")
parameters <- "x"
cmd <- paste(paste0("\"", executable, "\""), parameters, paste0("\"", file.path(path,  "Dataset.zip"), "\""))
system(cmd)
```

TThe file are copied in folder `UCI HAR Dataset` under the same working directpory. Lets list all the files here inside.

```{r}
pth <- file.path(pth, "UCI HAR Dataset")
list.files(pth, recursive=TRUE)
```


##Reading the files


####Read the subject files.

```{r}
dataSubjectTrain    <- fread(file.path(pth, "train", "subject_train.txt"))
dataSubjectTest  <- fread(file.path(pth, "test" , "subject_test.txt" ))
```

####Read the activity files.

```{r}
dataActivityTrain <- fread(file.path(pth, "train", "Y_train.txt"))
dataActivityTest  <- fread(file.path(pth, "test" , "Y_test.txt" ))
```

####Read the data files. 

```{r fileToDataTable}
fileToDataTable <- function (fileName) {
	df <- read.table(fileName)
	dt <- data.table(df)
}
dataTrain <- fileToDataTable(file.path(pth, "train", "X_train.txt"))
dataTest  <- fileToDataTable(file.path(pth, "test" , "X_test.txt" ))
```


##Read Data


####Concatenating the data tables.

```{r}
dataSubject <- rbind(dataSubjectTrain, dataSubjectTest)
setnames(dataSubject, "V1", "subject")
dataActivity <- rbind(dataActivityTrain, dataActivityTest)
setnames(dataActivity, "V1", "activityNum")
dt <- rbind(dataTrain, dataTest)
```

####Merging columns and setting key.

```{r}
dataSubject <- cbind(dataSubject, dataActivity)
dt <- cbind(dataSubject, dt)
setkey(dt, subject, activityNum)
```


##Mean and standard deviation


####For the mean and standard deviation, read the `features.txt` file

```{r}
dataFeatures <- fread(file.path(pth, "features.txt"))
setnames(dataFeatures, names(dataFeatures), c("featureNum", "featureName"))

dataFeatures <- dataFeatures[grepl("mean\\(\\)|std\\(\\)", featureName)]
```

####Creating a columns in DataFeature to be able to link to `dt`.

```{r}
dataFeatures$featureCode <- dataFeatures[, paste0("V", featureNum)]

```

####Subset these variables using variable names.

```{r}
select <- c(key(dt), dataFeatures$featureCode)
dt <- dt[, select, with=FALSE]
```


####Use descriptive activity names


Read `activity_labels.txt` file to set descriptive names to the activities.

```{r}
dataActivityNames <- fread(file.path(pth, "activity_labels.txt"))
setnames(dataActivityNames, names(dataActivityNames), c("activityNum", "activityName"))
```


Merging activity labels.

```{r}
dt <- merge(dt, dataActivityNames, by="activityNum", all.x=TRUE)
```

Add `activityName` as a key.

```{r}
setkey(dt, subject, activityNum, activityName)
```

Melt the data table to reshape it and adding new columns

```{r}
dt <- data.table(melt(dt, key(dt), variable.name="featureCode"))
dt <- merge(dt, dataFeatures[, list(featureNum, featureCode, featureName)], by="featureCode", all.x=TRUE)
dt$activity <- factor(dt$activityName)
dt$feature <- factor(dt$featureName)
```

Seperate features from `featureName` using the helper function `grepthis`.

```{r grepthis}
grepthis <- function (regex) {
  grepl(regex, dt$feature)
}
## Features with 2 categories

y <- matrix(seq(1, 2), nrow=2)
x <- matrix(c(grepthis("^t"), grepthis("^f")), ncol=nrow(y))
dt$featDomain <- factor(x %*% y, labels=c("Time", "Freq"))
x <- matrix(c(grepthis("Acc"), grepthis("Gyro")), ncol=nrow(y))
dt$featInstrument <- factor(x %*% y, labels=c("Accelerometer", "Gyroscope"))
x <- matrix(c(grepthis("BodyAcc"), grepthis("GravityAcc")), ncol=nrow(y))
dt$featAcceleration <- factor(x %*% y, labels=c(NA, "Body", "Gravity"))
x <- matrix(c(grepthis("mean()"), grepthis("std()")), ncol=nrow(y))
dt$featVariable <- factor(x %*% y, labels=c("Mean", "SD"))
## Features with 1 category
dt$featJerk <- factor(grepthis("Jerk"), labels=c(NA, "Jerk"))
dt$featMagnitude <- factor(grepthis("Mag"), labels=c(NA, "Magnitude"))
## Features with 3 categories

y <- matrix(seq(1, 3), nrow=3)
x <- matrix(c(grepthis("-X"), grepthis("-Y"), grepthis("-Z")), ncol=nrow(y))
dt$featAxis <- factor(x %*% y, labels=c(NA, "X", "Y", "Z"))
```

Check to make sure all possible combinations of `feature` are accounted for by all possible combinations of the factor class variables.

```{r}
r1 <- nrow(dt[, .N, by=c("feature")])
r2 <- nrow(dt[, .N, by=c("featDomain", "featAcceleration", "featInstrument", "featJerk", "featMagnitude", "featVariable", "featAxis")])
r1 == r2
```




###Create a tidy data set


Data set with the average of each variable for each activity-subject.

```{r}
setkey(dt, subject, activity, featDomain, featAcceleration, featInstrument, featJerk, featMagnitude, featVariable, featAxis)
dtTidy <- dt[, list(count = .N, average = mean(value)), by=key(dt)]
```

###Code Book

```{r}
save.image("myWorkSpace.RData")
knit("makeCodebook.Rmd", output = "codebook.md", encoding = "ISO8859-1")
markdownToHTML("codebook.md", "codebook.html")
```
